---
title: "Assignment 2"
author: "Isabel Hojman"
date: "14/10/2017"
output: html_document
---

## EXERCISE 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
source("functions.R")
```


## EXERCISE 1

a.
```{r}
print(sum_column(iris, "Sepal.Length"))
print(sum_column(iris, "Species"))
print(sum_column(warpbreaks, "breaks"))
```

b.
```{r}
print(my_sum(iris$Sepal.Length))
print(my_sum(iris$Species))
print(my_sum(warpbreaks$breaks))
```

c.
```{r}
print(sum_divided_by(iris$Sepal.Length, 12))
print(sum_divided_by(iris$Species, 22))
print(sum_divided_by(iris$Sepal.Length, "Not numeric"))
print(sum_divided_by(warpbreaks$breaks, -12))
```

d.
```{r}
print(my_mean(iris$Sepal.Length))
print(my_mean(iris$Species))
print(my_mean(warpbreaks$breaks))
```

## EXERCISE 2

a.
```{r}
print(grouped_violin_plot(iris, "Sepal.Length", "Species"))
```

b. 
```{r}
p <- grouped_violin_plot(iris, "Sepal.Length", "Species")
# YOUR CODE HERE: Change the colour scheme for the interior of the three violin plots
#                 to anything else at all.
p <- p + ggplot2::scale_fill_brewer(palette="PRGn")
# YOUR CODE HERE: Add a main title that says "Iris data".
p <- p + ggplot2::ggtitle("New Iris Data")
print(p)
```

## EXERCISE 3

a.
```{r}
difference_in_medians(iris, "Sepal.Width", "Species", "versicolor", "virginica")
difference_in_medians(iris, "Sepal.Width", "Species", "virginica", "virginica")
```

b. 
```{r}
iris$Sepal.Width[1:10]
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
randomize(iris, "Sepal.Width")$Sepal.Width[1:10]
randomize(iris, "Species")$Species[1:10]
randomize(iris, "Species")$Sepal.Width[1:10]
set.seed(previous_seed)
```

c.
```{r}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_1 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                 "virginica", difference_in_medians, n_samples=10)
ptest_2 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                 "virginica", difference_in_medians, n_samples=10)
ptest_3 <- permutation_twogroups(randomize(iris, "Sepal.Width"), "Sepal.Width",
                                 "Species", "versicolor", "virginica",
                                 difference_in_medians, n_samples=10)
set.seed(previous_seed)
print(ptest_1)
print(ptest_2)
print(ptest_3)
print(ptest_3$observed)
print(ptest_3[["observed"]])
```

d. Does it matter for your permutation test whether you permute var or grouping_var? Why or why not?
Yes, it matters. I have tested changing var to grouping_var and the result changes too. If we permute grouping_var, in the first example, for instance, what will be permute are the "Species" (setosa, versicolor, etc) and not the "Sepal.Width", that are the numbers that we want to permute.


e.
```{r, cache=T}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                              "virginica", difference_in_medians)
set.seed(previous_seed)
```

```{r}
ptest_d <- tibble::as_tibble(ptest["permuted"])
p <- ggplot2::ggplot(ptest_d, ggplot2::aes_string(x="permuted")) +
  ggplot2::geom_histogram(colour="black", fill="firebrick") +
  ggplot2::ylim(c(0, 6000)) +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest[["observed"]]), lwd=0.6, color="firebrick") +
  ggplot2::ggtitle("Difference in medians")
print(p)
```

What do you conclude from your plot?
We can see that the more frequent value for the difference of the medians in the permuted samples (0.0) is not the value observed for the entire samples (-0.2).

f.
```{r, cache=T}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_new <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                   "virginica", new_test_statistic)
set.seed(previous_seed)
```
```{r}
ptest_n <- tibble::as_tibble(ptest_new["permuted"])
p2 <- ggplot2::ggplot(ptest_n, ggplot2::aes_string(x="permuted")) +
  ggplot2::geom_histogram(colour="black", fill="yellow") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest_new[["observed"]]), lwd=0.6, color="firebrick") +
  ggplot2::ggtitle("Difference in means")
print(p2)
```

In the second histogram, we can see the difference betweens means in the permuted samples: the histogram shape is similar to the previous one, with a peack close to 0.0, but the samples are more distributed. The more frequent value is near 0.0, with almost 1000 samples; while in "Difference in medians", it is 0.0 with more than 5000 samples. In both cases, the observed value is -0.2 or close to it.

g.

```{r}
pvalue_r <- permutation_pvalue_right(ptest)
pvalue_l <- permutation_pvalue_left(ptest)
print(pvalue_r)
print(pvalue_l)
```

Let's take the example of A1 Spanish learners, divided in two groups: portuguese speakers and german speakers.
We'd predict that the difference in the medians statistic would be extreme on the right
side if our *Hypothesis B* were that the two groups are not equal in terms of time of response to an associate-image-to-word task.
In which case, we would use the function permutation_pvalue_right because if its value is extreme, the Hypotesis B would be considered true.

h.

```{r, cache=T}
iris_subset_1 <- iris[c(89:94, 108:112),]
iris_subset_2 <- iris[88:114,]

if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest1 <- permutation_twogroups(iris_subset_1, "Sepal.Width", "Species", "versicolor",
                              "virginica", difference_in_medians)
ptest2 <- permutation_twogroups(iris_subset_2, "Sepal.Width", "Species", "versicolor",
                              "virginica", difference_in_medians)
set.seed(previous_seed)
```

```{r}
ptest_1 <- tibble::as_tibble(ptest1["permuted"])
ptest_2 <- tibble::as_tibble(ptest2["permuted"])
p_1 <- ggplot2::ggplot(ptest_1, ggplot2::aes_string(x="permuted")) +
  ggplot2::geom_histogram(colour="black", fill="pink") +
  ggplot2::ylim(c(0, 6000)) +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest1[["observed"]]), lwd=0.6, color="pink") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest2[["observed"]]), lwd=0.6, color="blue") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest[["observed"]]), lwd=0.6, color="firebrick") +
  ggplot2::ggtitle("Difference in medians - Subset 1")
print(p_1)
p_2 <- ggplot2::ggplot(ptest_2, ggplot2::aes_string(x="permuted")) +
  ggplot2::geom_histogram(colour="black", fill="blue") +
  ggplot2::ylim(c(0, 6000)) +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest1[["observed"]]), lwd=0.6, color="pink") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest2[["observed"]]), lwd=0.6, color="blue") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest[["observed"]]), lwd=0.6, color="firebrick") +
  ggplot2::ggtitle("Difference in medians - Subset 2")
print(p_2)
```


* We can see that the three histograms are different. The one for the Subset 1 presents more samples in the values that are not so close to 0.00, while in the Subset 2 the samples are all around 0.00. In both cases, the  data is more sparse than in the histogram in 3e.

* The observed value for Subset 2 is closer to the observed value of the entire group than the one in the Subset 1, which is logical since the values for the permuted samples in Subset 2 are more similar to the orginal group values.

```{r}
print(permutation_pvalue_left(ptest1))
print(permutation_pvalue_right(ptest1))
print(permutation_pvalue_left(ptest2))
print(permutation_pvalue_right(ptest2))
```

* Based on the histogramms, it's reasonable to say that the groups of observations don't come from the same source of variability.
