---
title: "Assignment 2"
author: "Isabel Hojman"
date: "14/10/2017"
output: html_document
---

## EXERCISE 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
source("functions.R")
```


## EXERCISE 1

a.
```{r}
print(sum_column(iris, "Sepal.Length"))
print(sum_column(iris, "Species"))
print(sum_column(warpbreaks, "breaks"))
```

b.
```{r}
print(my_sum(iris$Sepal.Length))
print(my_sum(iris$Species))
print(my_sum(warpbreaks$breaks))
```

c.
```{r}
print(sum_divided_by(iris$Sepal.Length, 12))
print(sum_divided_by(iris$Species, 22))
print(sum_divided_by(iris$Sepal.Length, "Not numeric"))
print(sum_divided_by(warpbreaks$breaks, -12))
```

d.
```{r}
print(my_mean(iris$Sepal.Length))
print(my_mean(iris$Species))
print(my_mean(warpbreaks$breaks))
```

## EXERCISE 2

a.
```{r}
print(grouped_violin_plot(iris, "Sepal.Length", "Species"))
```

b. 
```{r}
p <- grouped_violin_plot(iris, "Sepal.Length", "Species")
# YOUR CODE HERE: Change the colour scheme for the interior of the three violin plots
#                 to anything else at all.
p <- p + ggplot2::scale_fill_brewer(palette="PRGn")
# YOUR CODE HERE: Add a main title that says "Iris data".
p <- p + ggplot2::ggtitle("New Iris Data")
print(p)
```

## EXERCISE 3

a.
```{r}
difference_in_medians(iris, "Sepal.Width", "Species", "versicolor", "virginica")
difference_in_medians(iris, "Sepal.Width", "Species", "virginica", "virginica")
```

b. 
```{r}
iris$Sepal.Width[1:10]
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
randomize(iris, "Sepal.Width")$Sepal.Width[1:10]
randomize(iris, "Species")$Species[1:10]
randomize(iris, "Species")$Sepal.Width[1:10]
set.seed(previous_seed)
```

c.
```{r}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_1 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                 "virginica", difference_in_medians, n_samples=10)
ptest_2 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                 "virginica", difference_in_medians, n_samples=10)
ptest_3 <- permutation_twogroups(randomize(iris, "Sepal.Width"), "Sepal.Width",
                                 "Species", "versicolor", "virginica",
                                 difference_in_medians, n_samples=10)
set.seed(previous_seed)
print(ptest_1)
print(ptest_2)
print(ptest_3)
print(ptest_3$observed)
print(ptest_3[["observed"]])
```

d. Does it matter for your permutation test whether you permute var or grouping_var? Why or why not?
Yes, it matters. I have tested changing var to grouping_var and the result changes too. If we permute grouping_var, in the first example, for instance, what will be permute are the "Species" (setosa, versicolor, etc) and not the "Sepal.Width", that are the numbers that we want to permute.


e.
```{r, cache=T}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                              "virginica", difference_in_medians)
set.seed(previous_seed)
```

```{r}
ptest_d <- tibble::as_tibble(ptest["permuted"])
p <- ggplot2::ggplot(ptest_d, ggplot2::aes_string(x="permuted")) +
  ggplot2::geom_histogram(colour="black", fill="orange") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest[["observed"]]), lwd=0.6, color="firebrick") +
  ggplot2::ggtitle("Difference in medians")
print(p)
```

What do you conclude from your plot?
We can see that the more frequent value for the difference of the medians in the permuted samples (0.0) is not the value observed for the entire samples (-0.2).

f.
```{r, cache=T}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_new <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                   "virginica", new_test_statistic)
set.seed(previous_seed)
```
```{r}
ptest_n <- tibble::as_tibble(ptest_new["permuted"])
p2 <- ggplot2::ggplot(ptest_n, ggplot2::aes_string(x="permuted")) +
  ggplot2::geom_histogram(colour="black", fill="yellow") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest_new[["observed"]]), lwd=0.6, color="firebrick") +
  ggplot2::ggtitle("Difference in means")
print(p2)
```

In the second histogram, we can see the difference betweens means in the permuted samples: the histogram shape is similar to the previous one, with a peack close to 0.0, but the samples are more distributed. The more frequent value is near 0.0, with almost 1000 samples; while in "Difference in medians", it is 0.0 with more than 5000 samples. In both cases, the observed value is -0.2 or close to it.

g.

h.
